---
title: "Koko Skills Test Analysis"
author: "Denis Munene"
date: " `r as.Date(Sys.time())` "
output:
  pdf_document:
    toc: yes
  html_document:
    number_sections: yes
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE, cache=F}

knitr::opts_chunk$set(echo = F, warning = F, message = F, fig.width = 10, fig.height = 10)

```

# Data Analysis


```{r packages}
## Importing Libraries

# load libraries

library(readxl)
library(openxlsx)
library(dplyr)
library(readxl)
library(janitor)
library(tidyr)
library(ggplot2)

```


## Importing Datasets and Preprocesing

```{r datasets, echo=TRUE}

# import datasets

dt1 <- readxl::read_xlsx("../AnalyticsSkillsTest-FictionalData.xlsx", sheet = "Sellers")
dt2 <- readxl::read_xlsx("../AnalyticsSkillsTest-FictionalData.xlsx", sheet = "Cooker_Sales")
dt3 <- readxl::read_xlsx("../AnalyticsSkillsTest-FictionalData.xlsx", sheet = "Fuel_Sales")


### rename id
dt2a <- dt2 %>% dplyr::rename(id = seller_id) 

### Join Sellers and Cookers dt
df1 <- dplyr::left_join(dt2a,dt1) ## by id


```


## Sales Productivity


### Question 1

### SQL Code

#### NOTE 

Use SQL Server Import and Export Wizard to Import the excel data on SQL or any other appropriate way before running codes below.\


##### CookersSold by Type\

use KokoData;
select type,  COUNT(distinct Cooker_Sales$.customer_id) as CookersSold
from Cooker_Sales$
left join Sellers$
on Cooker_Sales$.seller_id = Sellers$.id
group by type
order by CookersSold desc;

##### Fuel Sold by Customer Type and Territory\

select customer_type, sale_territory, SUM(litres_sold) as TotalLitersSold
from Cooker_Sales$
left join Sellers$
on Cooker_Sales$.seller_id = Sellers$.id
left join Fuel_Sales$
on Cooker_Sales$.customer_id = Fuel_Sales$.customer_id
group by customer_type, sale_territory
order by TotalLitersSold desc;

##### Fuel Sold by Seller Type\

select type, SUM(litres_sold) as TotalLitersSold
from Cooker_Sales$
left join Sellers$
on Cooker_Sales$.seller_id = Sellers$.id
left join Fuel_Sales$
on Cooker_Sales$.customer_id = Fuel_Sales$.customer_id
group by type
order by TotalLitersSold desc;

##### CookersSold by Type and Total Liters (Correct)\

select type, round(SUM(litres_sold),3) as TotalLitersSold, COUNT(distinct Cooker_Sales$.customer_id) as CookersSold
from Cooker_Sales$
left join Sellers$
on Cooker_Sales$.seller_id = Sellers$.id
left join Fuel_Sales$
on Fuel_Sales$.customer_id = Cooker_Sales$.customer_id
group by type
order by CookersSold desc;

### Plot1

```{r plot1, echo=FALSE}
#### CookersSold by Type

cooker_sales <-  df1 %>% group_by(type) %>% summarise(customer_id = n())

plot1 <- ggplot(data = cooker_sales, aes(y= customer_id, x = type,fill = type ))+ geom_bar(stat = "identity") +geom_text(aes(label=customer_id), vjust = 1.6, color = "black") + ggtitle("Cooker Sales Across Channels") +ylab("No of Customers") +theme_get()

plot1
```



```{r ,echo=FALSE}
#### Total Liters Purchased by Customer
dt3a <- dt3 %>% group_by(customer_id) %>% summarise(TotalLitresSold = sum(litres_sold))


### Fuel Sold by Seller Type
fuel_sales <- left_join(dt3,dt2a) ## by customer_id

fuel_sales_type <- left_join(fuel_sales,dt1) ## by id

#### Total Liters Sold by Seller type
litres_by_seller <- fuel_sales_type %>% group_by(type) %>% summarise(TotalLitresSold = sum(litres_sold))

```

### Plot2

```{r plot2, echo=FALSE}
### join with Fuel Sales
df2 <- left_join(dt3a,df1) ## by customer_id


#### Fuel Sold by Customer Type and Territory
litres_by_customer_terr <- df2 %>% group_by(customer_type,sale_territory) %>% summarise(TotalLitresSold = sum(round(TotalLitresSold)))


plot2 <- ggplot(data = litres_by_customer_terr, aes(y= TotalLitresSold, x = sale_territory, fill = customer_type ))+ geom_bar(position = "dodge", stat = "identity") + ggtitle("Liters Sold by Customer Type and Territory") + xlab("Territory")+coord_flip() +geom_text(aes(label=TotalLitresSold),position = position_dodge(.8),vjust = 1.1, color = "black")

plot2
```

### Plot3

```{r plot3, echo=FALSE}
#### Fuel Sold by Seller Type

plot3 <- ggplot(data = litres_by_seller, aes(y= TotalLitresSold, x = type, fill = type ))+ geom_bar(position = "dodge", stat = "identity") +geom_text(aes(label=TotalLitresSold),position = position_dodge(.8),vjust = 1.1, color = "black")+ ggtitle("Liters Sold by Seller Type") + xlab("type")

plot3
```


```{r table1, echo=FALSE}

table1 <- data.frame(litres_by_seller$type,litres_by_seller$TotalLitresSold)

knitr::kable(table1,"pipe", col.names = gsub("fuel_salesprod.","",names(table1)))

```


### Question 2

Referrer is most productive

### Question 3
Yes. Given that Refferers are making more cooker sales and simultaneously more fuel sales.


### Question 4
1) Agents making more fuel sales with less cooker sales compared to CSR hence more opportunity for productivity by increasing Agents
2) Focus on restaurants in North as well as households and restaurants in south initially as they appear more promising


### SQL Code


##### sales by seller type and Territory\

select type, sale_territory, round(SUM(litres_sold),3) as TotalLitersSold
from Cooker_Sales$
left join Sellers$
on Cooker_Sales$.seller_id = Sellers$.id
left join Fuel_Sales$
on Cooker_Sales$.customer_id = Fuel_Sales$.customer_id
group by type, sale_territory
order by TotalLitersSold desc;

### Plot4

```{r plot4, echo=FALSE}

#### Total Liters Sold by Seller type and territory
litres_by_seller_terr <- fuel_sales_type %>% group_by(type,sale_territory) %>% summarise(TotalLitresSold = sum(litres_sold))

#### Fuel Sold by Seller Type and territory

plot4 <- ggplot(data = litres_by_seller_terr, aes(y= TotalLitresSold, x = sale_territory, fill = type ))+ geom_bar(position = "dodge", stat = "identity") +geom_text(aes(label=TotalLitresSold),position = position_dodge(.8),vjust = 1.1, color = "black")+ ggtitle("Liters Sold by Seller Type and Territory") + xlab("sale_territory")

plot4

```


##### Sales by customer type and Territory for Agents\

select type, sale_territory,COUNT(distinct Cooker_Sales$.customer_id) as CookersSold, round(SUM(litres_sold),3) as TotalLitersSold
from Fuel_Sales$
left join Cooker_Sales$
on Fuel_Sales$.customer_id = Cooker_Sales$.customer_id
left join Sellers$
on Cooker_Sales$.seller_id = Sellers$.id
where type = 'Agent'
group by type, sale_territory
order by TotalLitersSold desc, CookersSold asc;

### Plot5

```{r plot5, echo=FALSE}

#### Total Liters Sold by Seller type and territory for Agents
litres_by_seller_terr_ag <- subset(litres_by_seller_terr,type =="Agent")

#### Fuel Sold by Seller Type and territory

plot5 <- ggplot(data = litres_by_seller_terr_ag, aes(y= TotalLitresSold, x = sale_territory, fill = type ))+ geom_bar(position = "dodge", stat = "identity") +geom_text(aes(label=TotalLitresSold),position = position_dodge(.8),vjust = 1.1, color = "black")+ ggtitle("Liters Sold by Seller Type and Territory for Agents") + xlab("sale_territory")

plot5

```



### Question 5

Price of cooker and fuel to assess price sensitivity


### Assumptions Made:
1) linear relationship btwn fuel consumption relative to number of cookers
2) same cost of cooker/fuel across locations
3) same price points across customer types
4) fair commission across seller types


## Fuel consumption


### Question 6


```{r ,echo=FALSE, results=FALSE}

#### Convert to date
str(fuel_sales_type$tx_date)
fuel_sales_type$tx_date <- as.Date(as.character(as.POSIXct(fuel_sales_type$tx_date)))

str(fuel_sales_type$tx_date)

### Convert to months

fuel_sales_type$tx_date <- months(fuel_sales_type$tx_date)


fuelconsumed_bymonth <- fuel_sales_type %>% group_by(type,tx_date) %>% summarise(Total_LitresSold = sum(litres_sold))

```


##### Total Monthly Fuel Consumption by Seller Type\

select type, MONTH (tx_date) as mwezi, round(SUM(litres_sold),3) as TotalLitersSold
from Cooker_Sales$
left join Sellers$
on Cooker_Sales$.seller_id = Sellers$.id
left join Fuel_Sales$
on Cooker_Sales$.customer_id = Fuel_Sales$.customer_id
group by type,MONTH (tx_date)
order by TotalLitersSold desc;



```{r table2, echo=FALSE}
table2 <- data.frame(fuelconsumed_bymonth$type,fuelconsumed_bymonth$tx_date,fuelconsumed_bymonth$Total_LitresSold)

knitr::kable(table2,"pipe", col.names = gsub("fuelconsumed_bymonth.","",names(table2)))
```
### Plot6

```{r}
#### Total Monthly Fuel Consumption by Seller Type

plot6 <- ggplot(data = fuelconsumed_bymonth, aes(y= Total_LitresSold, x = tx_date, fill = type ))+ geom_bar(position = "dodge", stat = "identity") +geom_text(aes(label=Total_LitresSold),position = position_dodge(.8),vjust = 1.1, color = "black")+ ggtitle("Total Monthly Fuel Consumption by Seller Type") + xlab("Month")

plot6



```


##### AvgFuel Sold per cooker by Seller Type and Month

select type,MONTH (tx_date) as mwezi, round(sum(litres_sold)/count(distinct Cooker_Sales$.customer_id),0) as average_literspercooker
from Cooker_Sales$
left join Sellers$
on Cooker_Sales$.seller_id = Sellers$.id
left join Fuel_Sales$
on Cooker_Sales$.customer_id = Fuel_Sales$.customer_id
group by type,MONTH (tx_date)
order by average_literspercooker desc;



```{r, echo=FALSE}

#### Convert to date

dt3$tx_date <- as.Date(as.character(as.POSIXct(dt3$tx_date)))

### Summarise Total litres sold to each customer

dt3$tx_date <- months(dt3$tx_date)

## Summarise Total litres sold to each customer
dt3b <- dt3 %>% group_by(customer_id,tx_date) %>% summarise(TotalLitresSold = sum(litres_sold))

dt3c <- left_join(dt3b, df1, by = "customer_id") %>% select(customer_id,tx_date,TotalLitresSold, type)

avg_fuelconsumed_bymonth <- dt3c %>% group_by(type,tx_date) %>% summarise(Total_LitresSold = sum(TotalLitresSold),CookersSold = n()) %>% mutate(average_literspercooker = round(Total_LitresSold/CookersSold))

avg_fuelconsumed_bymonth$tx_date <- factor(avg_fuelconsumed_bymonth$tx_date, levels = c("January","February","March"))


```





```{r table3, echo=FALSE}

table3 <- data.frame(avg_fuelconsumed_bymonth$type,avg_fuelconsumed_bymonth$tx_date,avg_fuelconsumed_bymonth$average_literspercooker)

knitr::kable(table3,"pipe", col.names = gsub("avg_fuelconsumed_bymonth.","",names(table2)))

```


### Plot7

```{r}

plot7 <- ggplot(data = avg_fuelconsumed_bymonth, aes(y= average_literspercooker, x = tx_date, fill = type ))+ geom_bar(position = "dodge", stat = "identity") +geom_text(aes(label=average_literspercooker),position = position_dodge(.9),vjust = 1.6, color = "black")+ ggtitle("Average monthly fuel consumption per cooker for each channel") + xlab("Month")

plot7
```



### Question 7

Agents produce most satisfied customers going by their month fuel consumption rates

### Question 8

1) Availability/closeness to customer hence more contact/personal/relatable with Agent 
2) More trust/reliability with Agent/refferee
3) Diversity of retail outlets at agent with respect t0 other items of sale (that support fuel consumption)

### Question 9
1) Focus on Agents and Refferals as channels
2) Subsidise prices i.e. discounts, scoring for bonus points/awards
3) Offer fuel for credit (on recommendation basis)
4) Support retail otluets marketing e.g through billboards
5) Product improvement & diferentiation


### Question 10
1) What's the rate of commission across channels?
2) Is the commission similar across territories? (maybe some territories deserve different incentive approach)
3) Avg (Size) and Composition of HH/restaurants across territories
4) Population size/ density of target regions

### Assumptions Made:
Similar fuel consumption across time (i.e. no cases of high demand/low supply) for the restaurants and households
